---
title: Regional variations in rates of total knee arthroplasty among Medicare beneficiaries
author: Michael M Ward and Abhijit Dasgupta
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->



<!-- badges: start -->
<!-- badges: end -->

This repository contains R code for some analyses and figure generation for our paper  looking at regional variations
in the rates of total knee arthroplasty across the United States among Medicare beneficiaries.

The paper is now published:

Ward, M. M., & Dasgupta, A. (2020). Regional Variation in Rates of Total Knee Arthroplasty Among Medicare Beneficiaries. JAMA network open, **3(4)**, e203717. [link](https://doi.org/10.1001/jamanetworkopen.2020.3717)

```{r, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F, cache = T,
                      fig.path = 'Result_Graphs/', dev ='png',
                      dpi = 300, 
                      # fig.align = 'center', 
                      fig.height = 5, fig.width = 7)

library(pacman)
pacman::p_load(char = c('tidyverse','sf','here','fs','rio'))
drop_dir <- path('P:/','Work','Ward','Studies','Medicare2015','data')
theme_set(theme_bw() + theme(
  text = element_text(size = 10),
  axis.title = element_text(size = 10, face = 'bold')))
map_scale_fill <- function(...){
  scale_fill_gradient2(
  name = 'OER',
  trans = 'log2',
  labels = function(x){round(x,2)},
  high = '#ff0000',
  low = '#003300',
  mid = '#ffff00',
  midpoint = 0,
  ...)
}
smr_white <- import(path(drop_dir, 'raw/PROJ4_SMR_WHITE.csv') )
hrr_info <- st_read(path(drop_dir, 'HRR_Bdry.SHP'), quiet = TRUE)
plot_dat <- smr_white %>% left_join(hrr_info, by = c('hrr' = 'HRRNUM'))# %>% 
  # mutate(SMR1 = ifelse(SMR1 <= 0.67, 0.67, SMR1)) %>% 
  # mutate(SMR1 = ifelse(SMR1 >= 1.5, 1.5, SMR1))
```
```{r, echo=FALSE}
plt11 <- 
  ggplot(plot_dat) +
  geom_sf(aes(geometry = geometry, fill = smr3),  color = NA) + 
  map_scale_fill(breaks = c(min(plot_dat$smr3),  1, max(plot_dat$smr3)))+
  # scale_fill_gradient2(
  #   name = 'OER',
  #   trans = 'log2',
  #   breaks = c(min(plot_dat$SMR3), 0.75, 1, 1.33, max(plot_dat$SMR3)),
  #   labels = function(x){round(x,2)},
  #   # limits = c(0.67, 1.5),
  #    high = 'red', mid = '#ffffbf', low = 'green', midpoint = 0) +
  coord_sf(label_graticule = 'SW', crs = 4286)

plt12 <- ggplotGrob(
  plt11 + coord_sf(xlim = c(-76,-71.5), ylim = c(39.5,42), crs=4286) + 
  theme(legend.position = 'none',
        axis.text = element_blank() , axis.ticks = element_blank())
)

plt13 <- plt11 + 
  
  annotation_custom(grob = plt12, xmin =-78.8, xmax = -63.8, ymin = 22.5, 
                    ymax = 32.5) +
  annotate('rect', xmax = -71.5, xmin = -76, ymax = 42, ymin = 39.5,
           alpha = 0.3)
print(plt13)
```

-----

This work was performed using `r sessioninfo::platform_info()$version` and the following additional
packages:

```{r}
pacman::p_load(char=c('tidyverse','fs'))
codefiles = fs::dir_ls(regex = 'R[md]*$', type = 'file', recurse=TRUE)
bl <- purrr::map(codefiles, ~grep('library\\(', readLines(.x), value=TRUE)) %>% 
  unlist() %>% 
  unique()
bl <- bl[stringr::str_detect(bl, '^#', negate=TRUE)]

pkgs <- stringr::str_match(bl, pattern='library\\(([:alnum:]+)\\)')[,2] %>% unique()
pacman::p_load(char=pkgs)

sessioninfo::package_info(pkgs) %>% filter(attached) %>%
  mutate(ghuser = str_match(source, 'Github \\(([:alnum:]+)')[,2]) %>%
  mutate(package = ifelse(is.na(ghuser), package, paste(ghuser, package, sep='/'))) %>%
  glue::glue_data("{package} ({loadedversion})") %>%
  paste(collapse='; ') %>%
  pander::pander()
```

> `<username>/<packagename>` refers to a package sourced from Github, that can be installed using `remotes::install_github("<username>/<packagename>")`
