---
title: The case for funnel plots
author: Abhijit
date: February 26, 2019
output:
    html_document:
        toc: true
        toc_float: true
        theme: journal
        highlight: zenburn
---

# Background

The statistical idea behind identifying certain HRR's as
having excess or deficit numbers of knee transplants after accounting
for demographic and clinical factors is to try and model the rate of
knee replacements in each region using Poisson regression (accounting
for repeated measures, if needed), and use this as indirectly
standardized rates accounting for these factors. We can then identify
regions which are outliers with respect to these standardized rates.

One point of confusion is in how to utilize these standardized rates.
The standardization can create a "normalized" distribution of rates,
and perhaps merely ranking the rates and seeing which regions change
rank the most suffices, or which regions move from "outliers" to
"normal" rates after adjustment works. However, this approach, in my
opinion, ignores natural variability in the estimates.

# A demonstrative simulation

We start by creating a population of 500,000 individuals split across
10 regions, , and
simulating realistic age, sex and race distributions for each region. The
regions differ in size and in local demographic distributions, and so the
rates have a natural variability solely dependent on the actual distribution
of race, age and sex within each region. Additionally, we added a protective
effect for regions 1 and 2 and an excess risk for regions 8, 9 and 10, which
represent non-demographic factors in each region that might drive the rate
of knee transplants. Ages were distributed as a normal distribution roughly
centered at 75 years and with a standard deviation of 5 years, the male: female
ratio was set roughly at 50:50 with some regional variability, and race, classified
as white, black and other, were simulated to have proportions in 45-60%,
10-30% for whites and blacks, with the rest being classified as other. The
following age, sex and race odds ratios were modeled:

```{r ors, echo = F, message = F, warning = F}
library(tidyverse)
tribble(~ Variable, ~ `Odds ratio`,
        'Age', '1.1 per year',
        'Male', '0.8',
        'Race: Black', '0,8',
        'Race: Other', '0.6',
        'Region: 1,2', '0.7',
        'Region: 8,9,10', '1.4') %>%
  knitr::kable()
```{r simulation_setup, echo = F, message = F, warning = F}
library(tidyverse)
library(broom)
set.seed(2095)
N <- 500000
reg_prob <- c(10,20, 5,15, 25, 10, 15, 5, 15,7)
reg_prob <- reg_prob/sum(reg_prob)
Region <- sample(1:10, N, replace = T,
                 prob = reg_prob)
dat <- tibble(Region)
dat <- dat %>% group_by(Region) %>%
  mutate(Age = rnorm(n(), rnorm(1, 75,3), rnorm(1, 5, 1))) %>%
  mutate(p = rbeta(1, 20, 20)) %>%
  mutate(Sex = sample(c('M','F'), n(), replace = T,
                      prob = c(1 - unique(p), unique(p)))) %>%
  ungroup()
dat$Race <- ""
for(i in 1:10){
  p_white <- runif(1, .45,.60)
  p_black <- runif(1, .10,.30)
  p_other <- 1 - p_white - p_black
  dat$Race[dat$Region == i] = sample(c('White','Black','Other'),
                                sum(dat$Region == i),
                                replace = TRUE,
                                prob = c(p_white, p_black, p_other))
}

linpred <- with(dat, log(0.005) + 0.1*(Age - 75) + log(0.8)*(Sex=='M') +
  log(1.4)*(Region >= 8) + log(0.7)*(Region <= 2) +
  log(0.8)*(Race=='Black') + log(0.6)*(Race == 'Other'))
p <- plogis(linpred)
outcome <- rbernoulli(N, p)
dat$outcome <- outcome
dat <- dat %>% mutate_if(is.character, as.factor) %>%
  mutate(Race = fct_relevel(Race, 'White','Black','Other'))
```

We then simulated the knee replacement statuses of 500,000 individuals,
with a baseline rate of 5 replacements per 1000 individuals. We summarize
the observed knee replacement rates by region in the bar graph below

```{r obs_rates, echo = F, message = F, warning = F}
dat %>% group_by(Region) %>% summarize(p_hat = mean(outcome)) %>%
  ggplot(aes(as.factor(Region), p_hat))+geom_bar(stat = 'identity') +
    labs(x = 'Region', y = 'Observed rates') +
  scale_y_continuous('Events per 10,000', labels = scales::unit_format(unit = '', scale = 1e4))
```

We note that, just by natural variability, region 6 has a higher rate of
replacements than regions 8, 9 or 10, which have an extraneous increased
rate of transplants in the simulation.

We then model the rates using GLMs (in this case logistic regression, but
Poission could have been used just as easily given the low baseline rate),
adjusting for age, race and sex. We also did nested models with just age, and with
age and sex as well. We'll concentrate on the full model, since it makes the point

```{r models, include = F, cache = TRUE}
m1 <- glm(outcome~Age,  data=dat, family = 'binomial')
m2 <- update(m1, . ~ . + Sex)
m3 <- update(m2, . ~ . + Race)

mods <- list('Model1' = m1, 'Model2' = m2, 'Model3' = m3)
checks <- map(mods, ~augment(., data = dat, type.predict = 'response') %>%
                group_by(Region) %>%
               summarize(obs = sum(outcome), expect = sum(.fitted),
                       SMR = obs/expect, N = n() ) %>%
                mutate(obs_rate = obs/N*10000, exp_rate = expect/N*10000))
```

We can then look at the predicted regional rate estimates from the model,
which can be considered as indirectly adjusted rates, compared to the observed
rates, in the boxplot below. The labels refer to the regions.

```{r boxplot, echo = F, warning = F, message = F}

bl <- checks[[3]] %>% select(Region, obs_rate, exp_rate) %>%
  rename(Observed = obs_rate, Expected = exp_rate) %>%
  gather(variable, value , -Region) %>%
  mutate(variable = forcats::fct_relevel(variable, 'Observed'))

ggplot(bl, aes(x = variable, y = value)) +
  geom_boxplot() +
  geom_label(aes(label = Region),
             position = position_jitter(width = 0.1)) +
  labs(x = '', y = 'Rate per 10,000') +
  theme_bw()
```

We see that regions 1, 8, 9, 10 fall in the box after adjustment,
which might be interpreted as being "normal" regions once you account
for demographics. Moreover, regions 4,5, 6, 7, which are "normal"
regions in the simulation, appear as outliers. This shows that
looking at the predicted (adjusted) rates, in not at all the right
approach.

# The case for statistical variability assessment and funnel plots

If we think a bit more about this quesiton, the null hypothesis we
are assuming is that there is no extraneous factors other than
demographics that explain the regional variabilty. If this is true,
then if we model the rates on the demographics, all that should be
left is statistical variability. Under this assumption of
residual statistical variability, we should look for outliers in
the "residuals" to see suggestions of the falsehood of our implied
null hypothesis. In Poisson regression, we are modeling the log-rates,
so the residuals can be considered as
$\log$(Observed) - $\log$(Expected), which is the logarithm of the
SMR. The standard funnel plot in this scenario plots the SMR against
the expected number of events, and puts control bounds based on
natural Poisson variabiity (or a normal approximation to it). We use
the normal approximation in the following plot, resulting from
modeling all the demographics,

```{r funnel, echo = F, warning = F, message = F}
funnel_SMR <- function(d){
  ggplot(d) + geom_label(aes(x = expect, y = SMR, label = Region)) +
    geom_line(aes(x = expect, y = 1 - 1.96 * sqrt(1/expect))) +
    geom_line(aes(x = expect, y = 1 + 1.96 * sqrt(1/expect))) +
    geom_line(aes(x = expect, y = 1 - 2.58 * sqrt(1/expect)), linetype = 2) +
    geom_line(aes(x = expect, y = 1 + 2.58 * sqrt(1/expect)), linetype = 2) +
    geom_hline(yintercept = 1, linetype = 2, color = 'darkgrey',
               size = 2) +
    theme_bw() +
    labs(x = 'Expected number of events', y = 'SMR')
}
plts <- map(checks, funnel_SMR)
plts[[3]]
```

We see in this plot that the correct regions are identified as being
at excess (8, 9, 10) and diminished (1,2) risk of replacements.

Let's expound on this a bit more. If the only factors driving the
replacement rate was demographics, we would expect the observed
rate to be close to the predicted rate when adjusting for age, race
and sex, while if there was some extraneous factor affecting the rate,
 we would see  poor agreement with the predicted rate, since the
 adjustment isn't accounting for an unmeasured factor.

# The fuller model

The underlying conceptual model that Mike proposes is

```
rate = [patient characteristics] + [physician characteristics] + [behavioral characteristics] + error
```

The part we're interested here is in the physician characteristics, with 
the behavioral characteristics being nuisance, in some sense. The physician characteristics are obtained as aggregate ecological measures like 

- Percentage with Medicare Advantage (`Mcare_adv_wtperc`)
- Density of orthopedic surgeions (`Ortho_per_100000`)

There are also some ecological measures for patient characteristics:

- % obesity (`obese_wtperc`)
- % with physical occupation (`physjob__wtperc`)
- % smokers (`smoking_wtperc`)

At the patient level we have 

- Age (continuous, `age_at_end_ref_yr`, discrete, `agecat`)
- Gender (`Male`, 1=male, 0=female)
- Race (`Race`, categorized as 1=white, 2=black, 3=hispanic, 4=asian, 5=other)
- A listing of 20 comorbidities, as 0/1 variables
- knee OA visits (continuous, `koa_visits`, discrete, `koa`, 0/1)
- knee symptom visits (continous, `kneesymptom_visits`, discrete, `kneesymptoms`, 0/1)
- Total knee visits (continuous, `allkneevisits`, discrete, `knee_patient`, 0/1)
- Dead or alive (`dead`, 1 = dead in year)
- Time of followup during calendar year (`personyrs`, 0-1 yrs)
- Current year of observation (`bene_enrollmt_ref_yr`, 2011-2015)
- Poor or on Medicaid (`poor`, 0/1)
