---
title: "Results of Statistical Analysis"
author: "Abhijit"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    toc: yes
  redoc::redoc:
    highlight_outputs: false
    comment_author: "Abhijit Dasgupta"
  pdf_document:
    toc: no
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=F, message = F, cache = F)
library(tidyverse)
library(rio)
theme_set(theme_bw() + theme(
  text = element_text(size = 10),
  axis.title = element_text(size = 10, face = 'bold')))
doc.type <- knitr::opts_knit$get('rmarkdown.pandoc.to')
print(doc.type)
```


## RESULTS

```{r Results-1, message = F, warning = F}
tka <- import('data/raw/TKAS2.csv') %>% 
  rename(year = BENE_ENROLLMT_REF_YR)
overall_smr <- import('data/raw/PROJ4_OVERALL_SMR.csv')
hrr_info <- sf::st_read('~/Downloads/hrr_bdry-1/HRR_Bdry.SHP',
                        quiet=TRUE)
hrr_info2 <- hrr_info %>% as.data.frame() %>% 
  select(HRRNUM, HRRCITY) %>% 
  separate(HRRCITY, c('State', 'City'), sep = '-', extra = 'merge') %>% 
  mutate(City = str_to_title(str_trim(City)))
overall <- overall_smr %>% select(hrr, SMR1, SMR3) %>% left_join(hrr_info2, by = c('hrr'='HRRNUM'))

```

In 2011, there were  `r format(tka$tka[tka$year==2011], big.mark=",")` TKA among
`r format(tka$n[tka$year==2011], big.mark=",")` beneficiaries in the study, while
in
2015, there were `r format(tka$tka[tka$year == 2015], big.mark=",")` TKA among `r format(tka$n[tka$year==2015], big.mark=",")` beneficiaries. When the expected rate of
TKA was based on adjustment for age, sex, and race/ethnicity, the OER
varied widely among HRRs (Supplemental figure 1-map). The highest OER of
`r overall %>% top_n(1, SMR1) %>% pull(SMR1) %>% round(digits=2)`  was in `r glue::glue_data(overall %>% top_n(1, SMR1), "{City}, {State}")`,  while the lowest OER of `r overall %>% top_n(1, -SMR1) %>% pull(SMR1) %>% round(digits=2)` was in 
`r glue::glue_data(overall %>% top_n(1, -SMR1), "{City}, {State}")`. 
HRRs with the highest OER were predominantly white, while HRRs with
the lowest OERs had large proportions of ethnic minorities (Supplemental
table 1). Despite adjustment of the expected rates for race/ethnicity,
significant correlations remained between the OER and the racial
composition of the HRR, indicating residual confounding (Supplemental
figure 2). Therefore, subsequent analyses used race-specific models to
generate the expected number of TKA. Because whites comprised 84.64% of the
sample, our analyses focused on associations among whites.

```{r Results-2}
poor <- import('data/raw/PROJ4_TKA_POOR.csv')
dementia <- import('data/raw/PROJ4_TKA_DEMENTIA.csv')

```


Among whites, the clinical characteristics of beneficiaries varied
widely among HRR, with for example, the percent of poor beneficiaries
ranging from `r round(100*min(poor$prop), 2)`% to 
`r round(100*max(poor$prop),2)`%, 
and those with dementia ranging from 
`r round(100*min(dementia$prop),2)`% to 
`r round(100*max(dementia$prop),2)`%. 
(Supplemental table 2). Adjustment for indicators of knee
osteoarthritis, comorbidities, and socioeconomic status resulted in OERs
that were somewhat less divergent, with 10^th^ and 90^th^ percentiles of
X and X, compared to X and X for OERs based on age and sex-adjustment
(Supplemental figure 3 and Supplemental table 3). However, substantial
regional variation in OERs remained after adjustment for patient
characteristics, with high OER in several HRR in the upper Midwest and
mountain west, and low OER in the New York City region and south Florida
(Figure 1 and Supplemental table 4).

HRRs that included more rural residents had generally higher OER than
those that were less rural (Figure 2). HRRs whose residents had fewer
outpatient visits for knee complaints also had higher OER than those
whose residents had more such visits. There was no association between
the OER and the proportion of Medicare Advantage beneficiaries in an
HRR. In contrast, HRRs with more TKA surgeons per capita had higher OERs
than those with fewer surgeons per capita.

HRRs with high OERs tended to have high OERs among patients with very
low estimated probabilities of TKA as well as those with higher
estimated probabilities of TKA, while HRRs with low OERs tended to have
low OERs across quartiles of estimated probability of TKA (Figure 3).
This pattern suggests that HRRs with high OERs were less discriminating
in performing TKA across a spectrum of beneficiaries with varying
likelihood of TKA, and that HRRs with low OERs were universally more
discriminating. Consistent with this interpretation, rates of TKA among
beneficiaries with dementia, peripheral vascular disease, and leg ulcers
were higher in HRRs with high OERs, as were rates among healthy 65 to 69
year-olds (Figure 4).

```{r phys}
phys <- read_csv('data/raw/phys_volume2.csv')
smr_white <- read_csv('data/raw/PROJ4_SMR_WHITE.csv')
phys <- phys %>% left_join(smr_white %>% select(hrr, smr3), by = c('hrrnum' = 'hrr'))
phys_per_hrr <- phys %>% count(hrrnum)
tka_per_phys <- phys %>% group_by(OP_PHYSN_NPI) %>% dplyr::summarize(N = sum(tka))
```

The number of TKA surgeons per HRR ranged from `r min(phys_per_hrr$n)` to 
`r max(phys_per_hrr$n)`. Most surgeons
regardless of HRR performed between X and X TKA annually (Figure 6).
Look for thresholds at 60.

```{r Results-3, echo = F}
smr_black <- read_csv('data/raw/Black_SMR.csv')
smr_black_filt <- smr_black %>% filter(n > 15000)
```

We limited the analysis of blacks and Hispanics to those HRRs that had
at least 15,000 black or Hispanic Medicare beneficiaries to provide
stable OER estimates. This number corresponded to the lowest HRR
population in the analysis of whites. Among blacks, there was
comparatively little variation in OERs among the `r sum(smr_black$n > 15000)`
HRRs studied, with a
range from `r round(min(smr_black_filt$smr1),2)` to 
`r round(max(smr_black_filt$smr1),2)` (Supplemental table 5). 

Among Hispanics, OERs ranged
from X to X among the X HRRs that were examined (Supplemental table 5).
HRRs with high OERs in whites also tended to have higher OERs among
blacks and Hispanics (Figure 5). However, there was no association
between TKA surgeon density in an HRR and the OER among blacks or
Hispanics (Figure 6).

## Figures

```{r Results-4, fig.width=7.5, cache = TRUE, eval=FALSE}
library(tidyverse)
library(rio)
library(sf)

smr_overall <- import('data/raw/PROJ4_OVERALL_SMR.csv') 
hrr_info <- st_read('~/Downloads/hrr_bdry-1/HRR_Bdry.SHP', quiet = TRUE)
plot_dat <- smr_overall %>% left_join(hrr_info, by = c('hrr' = 'HRRNUM'))# %>% 
  # mutate(SMR1 = ifelse(SMR1 <= 0.67, 0.67, SMR1)) %>% 
  # mutate(SMR1 = ifelse(SMR1 >= 1.5, 1.5, SMR1))

plt1 <- 
  ggplot(plot_dat) +
  geom_sf(aes(fill = SMR1)) + 
  scale_fill_gradient2(
    name = 'OER',
    trans = 'log',
    breaks = c(0.75, 1, 1.33 ),
    # limits = c(0.67, 1.5),
    low = 'orange', mid = '#ffffbf', high = 'green', midpoint = 0) +
  coord_sf(label_graticule = 'SW', crs = 4286)

# plt2 <- plt1 + coord_sf(xlim = c(-80, -70), ylim = c(40,45)) +
#   theme(legend.position = 'none',
#         axis.text = element_blank(),
#         axis.ticks = element_blank())

plt1
ggsave('graphs/map_model1.pdf', plot = plt1)
```

Figure 1.  Observed-to-expected ratios for rates of total knee arthroplasties
among white Medicare beneficiaries age 65 to 89 in 2011-2015, by Health Referral
Region.  Ratios greater than 1.0 indicate higher than expected rates of total
knee arthroplasty, while ratios less than 1.0 indicate lower than expected
rates. 

```{r Results-5, fig.width=7.5}
library(Hmisc)
library(patchwork)
smr_white <- import('data/raw/PROJ4_SMR_WHITE.csv')
label(smr_white$rural_perc) <- 'Percent rural\n'
label(smr_white$mean_op) <- 'Average number of \noutpatient visits per year'
label(smr_white$mcare_adv_wtperc) <- 'Percentage with \nMedicare Advantage'
label(smr_white$ortho_per_100000) <- 'Number of orthopedists \nper 100000 beneficiaries'
label(smr_white$mean_koa_visits) <- 'Average number of \nknee visits per person per year'
label(smr_white$frac_koa) <- 'Fraction of people with \nat least one knee visit a year'
plot_fn <- function(variable){
  vrbl <- enquo(variable)
  ggplot(smr_white, aes(x = !!vrbl, y = smr3))+
    geom_point()+
    geom_smooth(se = F, color = 'red') +
    geom_hline(yintercept = 1, linetype = 2) +
    labs(x = label(smr_white[[ensym(vrbl)]]),
         y = 'Observed/Expected ratio (OER)') +
    theme_bw() +
    theme(axis.text = element_text(size = 12),
          axis.title = element_text(size = 12, face = 'bold'),
          axis.title.y = element_blank())
}

p0 <- plot_fn(rural_perc)
p1 <- plot_fn(mean_op)
p2 <- plot_fn(mcare_adv_wtperc)
p3 <- plot_fn(ortho_per_100000)
p4 <- plot_fn(mean_koa_visits)
p5 <- plot_fn(frac_koa) + scale_x_continuous(labels=scales::percent_format())

plt <- ggpubr::ggarrange(p0,p1,p2,p3,p4,p5, ncol = 3, nrow = 2) 
(final_plt <- ggpubr::annotate_figure(plt, 
    left = ggpubr::text_grob('Observed/Expected ratio (OER)', 
                             rot = 90, size = 12, face = 'bold'))
)

```

Figure 2.  Associations between the percent of beneficiaries living in rural
areas, the number of outpatient visits for knee complaints, the percent of
beneficiaries in Medicare Advantage plans, and the number of surgeons performing
total knee arthroplasties per 10,000 beneficiaries in the and the
observed-to-expected ratio for rates of total knee arthroplasty among white
Medicare beneficiaries in each Health Referral Region.

```{r Results-6}
quarts <- rio::import('data/raw/PROJ4_MOD3_HRR_QUARTILES.csv') %>% as_tibble() %>%
  mutate(smr = tka/exp3) %>% select(hrr, smr, quartile)
smr_white <- rio::import('data/raw/PROJ4_SMR_WHITE.csv') %>% 
  select(hrr, smr3) %>% 
  mutate(quartile = 'Overall') %>% 
  mutate(smr_group = cut(smr3, quantile(smr3, c(0,0.05,0.4, 0.6, 0.95, 1)), 
                         include.lowest = T))
levels(smr_white$smr_group) <- str_split(levels(smr_white$smr_group), ',') %>% 
  map(~str_replace(., '\\(|\\)|\\[|\\]','')) %>% 
  map(as.numeric) %>% 
  map(~paste(round(., 2), collapse = '-')) %>% 
  unlist()
levels(smr_white$smr_group) <- paste(c('Very low','Low','Middle','High',
                                       'Very high'), 
                                     levels(smr_white$smr_group),
                                     sep = ': ')

blah <- quarts %>% bind_rows(smr_white %>% 
                               select(-smr_group) %>% 
                               rename(smr = smr3)) %>% 
  filter(!is.na(smr)) # Adding the overall SMR values to the dataset
blah <- blah %>% 
  left_join(smr_white %>% 
              select(hrr, smr_group)) %>%  # Adding the SMR groups to the dataset
  mutate(quartile = as_factor(quartile)) %>% 
  mutate(quartile = fct_recode(quartile,
                               "Low" = "q1",
                               "Low-middle" = "q2",
                               "High-middle" = "q3",
                               "High" = 'q4')) %>% 
  mutate(quartile = fct_relevel(quartile, 'Low','Low-middle', 'High-middle',
                                'High','Overall'))

ggplot(blah, aes(x = quartile,y = smr, color = smr_group)) +
  geom_line(aes(group = hrr), alpha = 0.7) + 
  geom_point() + 
  theme(legend.position = 'bottom',
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(nrow = 2, byrow = TRUE )) +
  labs(x = 'Quartile groups of patients based on expected probability of knee surgery',
       y = 'Probability quartile specific OER',
       color = 'HRR-specific overall OER quartiles') + 
  coord_cartesian(ylim = c(0,5))


```

Figure 3. Observed-to-expected ratios for rates of total knee arthroplasty in
each Health Referral Region among white Medicare beneficiaries, stratified by
beneficiary's expected probability of total knee arthroplasty.  Expected probabilities were
stratified into quartiles from very low (on the left) to highest (on the right),
and quartile-specific observed-to-expected ratios were computed for each region.

```{r Results-7}
conditions <- import('data/raw/PROJ4_TKA_CONDITIONS.csv')
conditions <- conditions %>% 
  filter(condition %in% c('dementia','pvd','ulcers','healthy')) %>% 
  mutate(condition = as.factor(condition)) %>% 
  mutate(condition = fct_recode(condition, 
                                'Dementia' = 'dementia',
                                'Peripheral vascular disease' = 'pvd',
                                'Leg ulcers' = 'ulcers',
                                'Healthy' = 'healthy'))
ggplot(conditions, aes(x = prop, y = smr3)) + 
  geom_point() + 
  geom_smooth(se = F, color = 'red') + 
  geom_hline(yintercept = 1, linetype = 2) +
  facet_wrap(~ condition, scales = 'free_x') +
  labs( x = 'Prevalence per 1000 subjects', 
        y = 'Observed/Expected Ratio (OER)') + 
  theme(text = element_text(size = 14))

```

Figure 4.  Associations between rates of total knee arthroplasty among white
Medicare beneficiaries with either dementia, peripheral vascular disease, leg
ulcers, or who were age 65 to 69 with no comorbidities and the
observed-to-expected ratio for rates of total knee arthroplasty by Health
Referral Region.


```{r}
library(rio)
conditions <- import('data/raw/PROJ4_TKA_CONDITIONS.csv')
conditions <- conditions %>% 
  filter(condition %in% c('dementia','pvd','ulcers','healthy')) %>% 
  mutate(condition = as.factor(condition)) %>% 
  mutate(condition = fct_recode(condition, 
                                'Dementia' = 'dementia',
                                'Peripheral vascular disease' = 'pvd',
                                'Leg ulcers' = 'ulcers',
                                'Healthy' = 'healthy'))
ortho_prevalence <- import('data/raw/PROJ4_SMR_WHITE.csv') %>% 
  select(hrr, ortho_per_100000)
plot_dat <- left_join(conditions, ortho_prevalence)
ggplot(plot_dat, aes(x = prop, y = ortho_per_100000)) + geom_point() + 
  geom_point() + 
  geom_smooth(se = F, color = 'red') + 
  facet_wrap(~ condition, scales = 'free_x') +
  labs( x = 'Prevalence per 1000 subjects', 
        y = 'Number of orthopedists per 100,000 beneficiaries') + 
  theme_bw()+
  theme(text = element_text(size = 14))
```

Figure 4a. Association between density of orthopedists and prevalence of demential pvd, leg ulcers or healthy individuals among
white Medicare beneficiaries by HRR

```{r Results-8}
smr_black <- read_csv('data/raw/Black_SMR.csv') %>% 
  filter(n >= 15000)
smr_white <- read_csv('data/raw/PROJ4_SMR_WHITE.csv')

smrs <- bind_rows(smr_white, smr_black) %>% 
  mutate(race = ifelse(race == 1, 'White','Black')) %>%
  mutate(race = as_factor(race)) %>% 
  # mutate(race = fct_relevel('White')) %>% 
  select(hrr, race, smr3, ortho_per_100000)

smrs %>% select(-ortho_per_100000) %>% 
  spread(race, smr3) %>% 
  ggplot(aes(White, Black)) + 
  geom_point() + 
  geom_smooth(se=F, color = 'red', method = 'lm') + 
  geom_abline(linetype = 2) + 
  coord_equal(xlim = c(0.6, 1.3), ylim = c(0.6, 1.3)) + 
  theme_bw() + 
  theme(text = element_text(size = 12),
        axis.title = element_text(face = 'bold'))
# ggplot(smrs, aes(x = ortho_per_100000, y = smr3, group = race, color = race)) +
#   geom_point(alpha = 0.5) + geom_smooth(se = F) +
#   # scale_y_log10() +
#   geom_hline(yintercept = 1, linetype = 2) +
#   # scale_color_manual(values = c('White'='black','Black' = 'blue')) + 
#   labs(x = 'Number of surgeons performing TKA\nper 10,000 beneficiaries',
#        y = 'Observed/Expected ratio (OER)',
#        color = 'Race') + 
#   theme(text = element_text(size = 14))
```


Figure 5.  Association between the observed-to-expected ratios for rates of
total knee arthroplasty between white and black 
Medicare beneficiaries.  Associations between the observed-to-expected ratios
for rates of total knee arthroplasty in white and black Medicare
beneficiaries and the number of surgeons performing total knee arthroplasties
per 10,000 beneficiaries in the Health Referral Region. HRRs for blacks are restricted to those with at least 15,000
black beneficiaries. 


```{r}
smr_black <- read_csv('data/raw/Black_SMR.csv') %>% 
  filter(n >= 15000)
ggplot(smr_black, aes(x = ortho_per_100000, y = smr3)) +
  geom_point() + geom_smooth(color = 'red', se = F) + 
  geom_hline(yintercept = 1, linetype = 2) +
  labs(x = 'Number of orthopedists \nper 100000 beneficiaries',
       y = 'Observed/Expected ratio (OER)')
```

Figure 6.  Association of the OER among blacks with the number of orthopedists 
per 100000 beneficiaries in a HRR. 

## SUPPLEMENTAL TABLES & FIGURES

```{r Results-9}
library(kableExtra)
library(sf)
library(gt)
library(flextable)
overall_smr <- read_csv('data/raw/PROJ4_OVERALL_SMR.csv')
hrr_info <- st_read('~/Downloads/hrr_bdry-1/HRR_Bdry.SHP', quiet = TRUE) %>%
  as_tibble() %>%
  select(HRRNUM, HRRCITY) %>% 
  separate(HRRCITY, c('State','City'), sep = '-', extra = 'merge') %>% 
  mutate(City = str_to_title(str_trim(City)))

top10 <- overall_smr %>% 
  arrange(SMR1) %>% 
  slice(1:10, (n() - 9):n())
top10 <- top10 %>% 
  left_join(hrr_info, by = c('hrr' = 'HRRNUM'))

races <- read_csv('data/raw/PROJ4_RACE.csv')
races <- races %>% spread(race, Percent) %>%
  set_names(c('HRR','White','Black','Hispanic','Asian','Other'))
top10 <- top10 %>% left_join(races, by = c('hrr' = 'HRR')) %>%
  select(hrr, SMR1, City, State, White:Other) %>% 
  mutate_at(vars(White:Other), ~./100)

if(doc.type == 'pdf'){
gt(top10) %>%
  tab_row_group(group = "Lowest 10", rows = 1:10) %>%
  tab_row_group(group = 'Highest 10', rows = 11:20) %>%
  fmt_number(vars(SMR1), decimals = 2) %>%
  fmt_percent(vars(White,Black,Hispanic,Asian,Other)) %>%
  cols_label(hrr = "HRR",
             SMR1 = "OER")
} else {
top10 <- top10 %>%  mutate_at(vars(White:Other), ~.*100) %>% 
  add_row(hrr='Lowest 10', .before = 1) %>% 
  add_row(hrr='Highest 10', .after = 11)
  
ft <- flextable(top10)
ft <- set_header_labels(ft, SMR1 = 'OER', hrr='HRR')
ft <- colformat_int(x = ft, col_keys = c('hrr'))
ft <- colformat_num(x = ft, col_keys = c('White','Black','Hispanic','Asian','Other'),
                    digits=2, suffix = "%")
fp1 <- officer::fp_border()
ft <- hline(ft, i = c(1, 11, 12,22), border = fp1)
ft <- bold(ft, part = 'header')
ft
}
```

__Supplemental Table 1:__ Racial composition in the HRRs with the highest and 
lowest OER when adjusted for age, race and sex

```{r Results-10}
suppressPackageStartupMessages(library(tidyverse))
library(gt)
library(flextable)
if(doc.type != 'pdf'){
suppl2 <- rio::import('data/raw/SupplTab2_2.csv')
out <- suppl2 %>% select(-hrr) %>% 
  gather(variable, value) %>% 
  group_by(variable) %>% 
  dplyr::summarize(Min = min(value), `10%` = quantile(value, 0.1),
            Median = median(value),
            `90%` = quantile(value, 0.9),
            Max = max(value))
ind <- which(out$Max >=1)
ind1 <- setdiff(1:nrow(out), ind)
out1 <- out %>% mutate_at(vars(Min:Max),
                          ~ifelse(out$Max >= 1, ., 100*.))
colvars <- out %>% select(Min:Max) %>% names()

grp1 <- c('age','race')
grp3 <- c('kneesymptoms','knee_patient','op_visits')
grp4 <- c('obese','physjob','smoking')
grp5 <- c('mcare','ortho','poor','zscore','rural')
grp2 <- setdiff(out1$variable, c(grp1,grp3,grp4,grp5))
grps <- c(grp1, grp2, grp3, grp4, grp5)

varlabs <- c('age'='Age', 'race' = 'Percent white', 'acutemi' = 'Acute MI', 
             'afib' = 'Atrial fibrilation', 'breast' = 'Breast Cancer','chf' = 'Chronic heart failure',
             'ckd' = 'Chronic kidney disease', 'colorectal' = 'Colorectal cancer', 
             'copd' = 'COPD', 'dementia' = 'Dementia', 'depress' = 'Depression',
             'diab' = 'Diabetes mellitus', 'endometrial_past' = 'Endometrial cancer', 
             'hememalignancy'='Heme malignancy', 'hiv' = 'HIV', 'ihd' = 'Ischemic heart disease',
             'koa' = 'Any knee OA visit')

out1 <- out1[match(grps, out1$variable ),]

ft1 <- flextable(out1) %>% 
  colformat_num(col_keys = colvars, 
                digits = 2) %>% 
  colformat_num(col_keys = colvars, 
                i = ind1,
                suffix = '%') %>% 
  bold(part = 'header')
ft1
} else {

gt(out) %>%
  fmt_percent(columns = 2:6, rows = (Max <= 1)) %>%
  fmt_number(columns = 2:6, rows = (Max > 1), decimal = 2)
}
```


__Supplemental Table 2:__ Distribution of demographic and clinical characteristics across HRRs

```{r Results-11, echo = F }
library(tidyverse)

overall <- read_csv('data/raw/PROJ4_OVERALL_SMR.csv')
races <- read_csv('data/raw/PROJ4_RACE.csv')
races <- races %>% spread(race, Percent) %>%
  set_names(c('HRR','White','Black','Hispanic','Asian','Other'))

overall2 <- overall %>% left_join(races, by = c('hrr' = 'HRR')) %>% 
  mutate_at(vars(White:Other), ~./100) %>% 
  select(-Other) %>% 
  gather(race, value, White:Asian) %>% 
  mutate(race = as_factor(race)) 

plt <- ggplot(overall2, aes(x = value, y = SMR1)) +
  geom_point() + geom_smooth(color = 'red', se = F) +
  facet_wrap(~race, scales = 'free_x') + 
  geom_hline(yintercept = 1, linetype = 2) + 
  scale_x_continuous(name = 'Percentage in a HRR', 
                     labels = scales::percent_format()) + 
  labs(y = 'Observed/Expected ratio (OER)') +
  theme_bw() + 
  theme(text = element_text(size = 12),
        axis.title = element_text(face = 'bold'))
plot(plt)
```

__Supplemental Figure 2:__ Association of OER based on age, race and sex adjustment with racial composition in the HRR




__Supplemental Table 5:__ 
