---
title: "Results of Statistical Analysis"
author: "Abhijit"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  word_document:
    reference_docx: ref_pgbrk.docx
  redoc::redoc:
    highlight_outputs: false
    comment_author: "Abhijit Dasgupta"
  pdf_document:
    toc: no
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F, cache = F,
                      fig.path = 'Result_Graphs/', dev = c('png','pdf'),
                      dpi = 300, 
                      fig.align = 'center', fig.height = 5, fig.width = 7)
library(tidyverse)
library(rio)
theme_set(theme_bw() + theme(
  text = element_text(size = 10),
  axis.title = element_text(size = 10, face = 'bold')))
doc.type <- knitr::opts_knit$get('rmarkdown.pandoc.to')

library(fs)
#drop_dir <- path(ProjTemplate::find_dropbox(), 'NIAMS','Ward','Medicare2015')
drop_dir <- path('P:/','Work','Ward','Studies','Medicare2015','data')
print(doc.type)
```


## RESULTS

```{r Results-1, message = F, warning = F}
tka <- import(path(drop_dir,'raw/TKAS2.csv')) %>% 
  rename(year = BENE_ENROLLMT_REF_YR)
overall_smr <- import(path(drop_dir,'raw/PROJ4_OVERALL_SMR.csv'))
hrr_info <- sf::st_read(path(drop_dir, 'HRR_Bdry.SHP'),
                        quiet=TRUE)
hrr_info2 <- hrr_info %>% as.data.frame() %>% 
  select(HRRNUM, HRRCITY) %>% 
  separate(HRRCITY, c('State', 'City'), sep = '-', extra = 'merge') %>% 
  mutate(City = str_to_title(str_trim(City)))
overall <- overall_smr %>% select(hrr, SMR1, SMR3) %>% left_join(hrr_info2, by = c('hrr'='HRRNUM'))

```

In 2011, there were  `r format(tka$tka[tka$year==2011], big.mark=",")` TKA among
`r format(tka$n[tka$year==2011], big.mark=",")` beneficiaries in the study, while
in
2015, there were `r format(tka$tka[tka$year == 2015], big.mark=",")` TKA among `r format(tka$n[tka$year==2015], big.mark=",")` beneficiaries. When the expected rate of
TKA was based on adjustment for age, sex, and race/ethnicity, the OER
varied widely among HRRs (Supplemental figure 1-map). The highest OER of
`r overall %>% top_n(1, SMR1) %>% pull(SMR1) %>% round(digits=2)`  was in `r glue::glue_data(overall %>% top_n(1, SMR1), "{City}, {State}")`,  while the lowest OER of `r overall %>% top_n(1, -SMR1) %>% pull(SMR1) %>% round(digits=2)` was in 
`r glue::glue_data(overall %>% top_n(1, -SMR1), "{City}, {State}")`. 
HRRs with the highest OER were predominantly white, while HRRs with
the lowest OERs had large proportions of ethnic minorities (Supplemental
table 1). Despite adjustment of the expected rates for race/ethnicity,
significant correlations remained between the OER and the racial
composition of the HRR, indicating residual confounding (Supplemental
figure 2). Therefore, subsequent analyses used race-specific models to
generate the expected number of TKA. Because whites comprised 84.64% of the
sample, our analyses focused on associations among whites.

```{r Results-2}
poor <- import(path(drop_dir,'raw/PROJ4_TKA_POOR.csv'))
dementia <- import(path(drop_dir, 'raw/PROJ4_TKA_DEMENTIA.csv'))

```


Among whites, the clinical characteristics of beneficiaries varied
widely among HRR, with for example, the percent of poor beneficiaries
ranging from `r round(100*min(poor$prop), 2)`% to 
`r round(100*max(poor$prop),2)`%, 
and those with dementia ranging from 
`r round(100*min(dementia$prop),2)`% to 
`r round(100*max(dementia$prop),2)`%. 
(Supplemental table 2). Adjustment for indicators of knee
osteoarthritis, comorbidities, and socioeconomic status resulted in OERs
that were somewhat less divergent, with 10^th^ and 90^th^ percentiles of
X and X, compared to X and X for OERs based on age and sex-adjustment
(Supplemental figure 3 and Supplemental table 3). However, substantial
regional variation in OERs remained after adjustment for patient
characteristics, with high OER in several HRR in the upper Midwest and
mountain west, and low OER in the New York City region and south Florida
(Figure 1 and Supplemental table 4).

HRRs that included more rural residents had generally higher OER than
those that were less rural (Figure 2). HRRs whose residents had fewer
outpatient visits for knee complaints also had higher OER than those
whose residents had more such visits. There was no association between
the OER and the proportion of Medicare Advantage beneficiaries in an
HRR. In contrast, HRRs with more TKA surgeons per capita had higher OERs
than those with fewer surgeons per capita.

HRRs with high OERs tended to have high OERs among patients with very
low estimated probabilities of TKA as well as those with higher
estimated probabilities of TKA, while HRRs with low OERs tended to have
low OERs across quartiles of estimated probability of TKA (Figure 3).
This pattern suggests that HRRs with high OERs were less discriminating
in performing TKA across a spectrum of beneficiaries with varying
likelihood of TKA, and that HRRs with low OERs were universally more
discriminating. Consistent with this interpretation, rates of TKA among
beneficiaries with dementia, peripheral vascular disease, and leg ulcers
were higher in HRRs with high OERs, as were rates among healthy 65 to 69
year-olds (Figure 4).

```{r phys}
phys <- read_csv(path(drop_dir,'raw/phys_volume2.csv'))
smr_white <- read_csv(path(drop_dir, 'raw/PROJ4_SMR_WHITE.csv'))
phys <- phys %>% left_join(smr_white %>% select(hrr, smr3), by = c('hrrnum' = 'hrr'))
phys_per_hrr <- phys %>% count(hrrnum)
tka_per_phys <- phys %>% group_by(OP_PHYSN_NPI) %>% dplyr::summarize(N = sum(tka))
```

The number of TKA surgeons per HRR ranged from `r min(phys_per_hrr$n)` to 
`r max(phys_per_hrr$n)`.  The annual number of TKA performed varied widely, but
the number and range of TKAs per surgeon was similar in high OER regions as in
other regions (Supplemental figure 4).
```{r, include=F}
smr_black_filt <- import(path(drop_dir,'Black_SMR.csv')) %>% 
  filter(n > 15000)
```

Analyses in blacks, Hispanics, and Asians was limited by low representation
across HRRs.  Only six HRR included at least 15,000 Hispanic beneficiaries,
which corresponded to the lowest white HRR population in our study.  Forty-two
HRRs included at least 15,000 black beneficiaries.  OERs among blacks in these
HRRs ranged from `r round(min(smr_black_filt$smr1),2)` to 
r round(max(smr_black_filt$smr1),2)` (Supplemental table 5).  OERs in blacks and whites
in these regions were positively correlated, and generally higher among blacks
(Figure 5).


## Hawaii

The Hawaii HRR information 

```{r hawaii}
library(tidyverse)
library(rio)
data_dir <- fs::path('P:/','Work','Ward','Studies','Medicare2015')
smr_white <- import(path(data_dir, 'data/raw/PROJ4_SMR_WHITE.csv'))
smr_overall <- import(path(data_dir, 'data/raw/PROJ4_OVERALL_SMR.csv'))
hrr_info <- sf::st_read(fs::path(data_dir, 'data','HRR_Bdry.SHP'), quiet = TRUE)

smr_hawaii <- smr_overall %>% left_join(hrr_info, by = c('hrr'='HRRNUM')) %>% 
  select(hrr, HRRCITY, SMR1, SMR3) %>% 
  filter(hrr == 150) %>% 
  separate(HRRCITY,c('State','City'), sep='-') %>% 
  select(hrr, City, State, everything())

library(flextable)
flextable(smr_hawaii) %>% 
  set_header_labels(hrr='HRR', SMR1='OER (Model 1)', SMR3 = 'OER (Model 3)') %>% 
  theme_box()

```



#### Table 1.  Observed/expected ratios (OER) for primary total knee arthroplasty rates among white Medicare beneficiaries in the urban centers or outlying areas of selected Health Referral Regions.  Regions selected were those with a large geographic area, an urban center with a population of 100,000 or more, and an OER at the low, middle, or high end of the distribution.

```{r Table-1}
d <- read_csv(path(drop_dir, 'raw/PROJ4_SMR_RURAL_URBAN.csv'))
hrr_info <- sf::st_read(path(drop_dir, 'HRR_Bdry.SHP'), quiet=T) %>% 
  as_tibble() %>% 
  select(HRRNUM, HRRCITY, -geometry) %>% 
  separate(HRRCITY, c('State','City'), sep = '-', extra='merge') %>% 
  mutate(City = str_to_title(str_trim(City))) %>% 
  unite('Location', c('City','State'), sep = ', ')

d1 <- d %>% mutate(city = ifelse(city == 0, 'Rural','Urban')) %>%
  select(hrrnum, city, smr3, location) %>%
  spread(city, smr3) %>%
  mutate(Difference = Rural - Urban) %>%
  arrange(Urban)


d2 <- d %>% select(hrrnum, overall_smr) %>% distinct()
d3 <- d1 %>% left_join(d2) %>%
  arrange(overall_smr) %>%
  rename('Overall OER' = 'overall_smr')

d4 <- d %>% filter(city==1) %>% select(hrrnum, Percent)
d3 <- d3 %>% left_join(d4) %>% rename('Percent Urban'= 'Percent')

d3 <- d3 %>% 
  left_join(hrr_info, by = c('hrrnum'='HRRNUM')) %>% 
  select(Location, `Percent Urban`, `Overall OER`, Urban, Rural, Difference, -location, -hrrnum) %>% 
  mutate(`Percent Urban` = as.numeric(str_replace(`Percent Urban`, '%','')))

ft <- flextable(d3) %>% 
  # theme_box() %>% 
  colformat_num(col_keys = names(d3)[3:6],
                digits = 2) %>% 
  colformat_num(col_keys=c('Percent Urban'), digits=1) %>% 
  align(j = 2:5, align='center', part='all') %>% 
  set_header_labels(values = c("Percent Urban" = "Percent of \nbeneficiaries living in\n the urban center",
                               "Urban" = 'OER urban center',
                               "Rural" = 'OER outlying area',
                               'Difference' = 'OER difference',
                               'Location' = 'Heath Referral\n Region')) %>% 
  autofit()
ft
```

