---
title: Updated analyses for journal response
author: Abhijit Dasgupta
date: "`r format(Sys.time(), '%d %b, %Y %I:%M %p')`"
output_format: html_document
editor_options: 
  chunk_output_type: console
---

```{r preamble, include = F}
knitr::opts_chunk$set(warning = F, message = F, echo = F, cache=TRUE)
library(tidyverse)
library(fs)
library(kableExtra)
library(rio)
library(here)
library(sf)
library(infer)
library(sjlabelled)

if(Sys.info()['sysname']=='Windows'){
  data_dir <- path('P:/','Work','Ward','Studies','Medicare2015','data')
  dir_exists(data_dir)
}
```


## White demographics

> For whites only, need number of beneficiaries and number of TKA, by year and overall.  Among these, need descriptives for age and sex by year and overall, and percent with knee osteoarthritis or knee symptoms. Also crude rate of TKA overall among whites. (E-mail, 2/4/2020 1:43 PM)

```{r, eval=F}
whites_tka <- import(path(data_dir,'raw','TKAS_WHITE2.csv')) %>% 
  select(BENE_ENROLLMT_REF_YR, tka, n) %>%
  # janitor::adorn_totals('row') %>% 
  mutate(tka = format(tka, big.mark=','), n = format(n, big.mark=',')) %>%
  rename('Year'='BENE_ENROLLMT_REF_YR','TKA'='tka', 'Number of beneficiaries' = 'n')
# kable(whites_tka) %>% 
#   kable_styling(full_width=F)
kable(whites_tka, align='lcc',
      caption = "Table 1: Number of TKA and number of beneficiaries among whites by year")
# pander::pander(whites_tka,
#                caption = "Table 1: Number of TKA and number of beneficiaries among whites by year")
```

```{r}
whites <- import(path('P:/', 'Work','Ward','Studies','Medicare2015','docs','revision_updates.xlsx'), 
                 n_max = 6) %>% 
  mutate(fraction_male = 100*(1-fraction_male)) %>% 
  # select(Year, MeanAge:perc_knee) %>% 
  mutate_at(vars(TKA:`Number Beneficiaries`), format, big.mark=',') %>% 
  mutate_at(vars(MeanAge:perc_knee), round, 2) %>% 
  rename('Age (mean)' = 'MeanAge', 'Age (SD)' = 'SDAge', 'Percent female' = 'fraction_male', 'Percent with knee issues' = 'perc_knee')
kable(whites, align = 'lcccccc',caption = "Table 1: TKA, benficiaries, age, sex and knee issues by year") %>% 
  kable_styling(full_width=T)
```

```{r}
smr_white <- import(path(data_dir, 'raw', 'PROJ4_SMR_WHITE.csv'))
```

The crude rate for TKA is `r round(1000*sum(smr_white$total_knee)/sum(smr_white$py),2)` per 1000 person-years. 


## Correlation analysis

> They also are asking for measures of effect size and variation.  I am thinking we should add correlations between OER and the predictors in the figures.  What do you think? (E-mail, 2/4/2020 2:03 PM)

### Figure 2

Correlation of various HRR-based summary measures with SMR along with bootstrap (n=5000)-based confidence intervals 

```{r}
library(mosaic)

smr_white <- import(path(data_dir, 'raw/PROJ4_SMR_WHITE.csv'))

dat <- smr_white %>% select(hrr, smr3, mean_op, rural_perc:mean_koa_visits, frac_koa) %>% 
  as_tibble()

cor_obs <- dat %>%
  pivot_longer(cols = mean_op:frac_koa, names_to="metric", values_to="frac") %>% 
  group_by(metric) %>% 
  summarize(corr_coef = cor(smr3, frac, method='spearman'))

set.seed(20294)
set.rseed(20294)
boot_samples <- do(5000, parallel=FALSE)*(resample(dat) %>% 
                           pivot_longer(cols=mean_op:frac_koa, names_to='metric',
                                        values_to='frac') %>% 
                           group_by(metric) %>% 
                           summarize(corr_coef = cor(smr3, frac,
                                                     method='spearman')) %>% 
                           ungroup())
cor_boot <- boot_samples %>% 
  group_by(metric) %>% 
  summarize(lcb = quantile(corr_coef, 0.025),
            ucb = quantile(corr_coef, 0.975))

labs <- c(
  'rural_perc' = 'Percent rural',
  'mean_op' = 'Average number of outpatient visits per year',
  'mcare_adv_wtperc' = 'Percentage with Medicare Managed Care',
  'ortho_per_100000' = 'Number of TKA surgeons per 100,000 beneficiaries',
  'mean_koa_visits' = 'Mean annual knee visits per person',
  'frac_koa' = 'Fraction of people with at least one knee visit a year'
)
out <- cor_obs %>% left_join(cor_boot) %>% 
  mutate_at(vars(corr_coef:ucb), round, 3) %>% 
  mutate(metric = labs[metric]) %>% 
  slice(c(6,3, 4, 2, 5))
kable(out, 
      col.names = c('Measure', 'Correlation','LCB','UCB')) %>% 
  kable_styling(full_width = TRUE)
```

### Figure 4

We computed the Spearman correlation coefficient between the prevalences and the corresponding SMRs and computed a 95% confidence interval using 5000 bootstrap samples. The results are below.

```{r, eval=T}
library(mosaic)

drop_dir <- data_dir
conditions <- import(path(drop_dir, 'raw/PROJ4_TKA_CONDITIONS_june2.csv'))
depr <- import(path(drop_dir, 'raw/PROJ4_TKA_DEPRESSION.csv')) %>% 
  mutate(condition='depression') %>% mutate(prop = prop*1000)
conditions <- bind_rows(conditions,depr)
conditions <- conditions %>% 
  filter(condition %in% c('dementia','ulcers','pvd','chf','depression','diabetes','healthy')) %>% 
  mutate(condition = as.factor(condition)) %>% 
  mutate(condition = fct_relevel(condition, 'dementia','ulcers','pvd','chf','depression','diabetes','healthy')) %>% 
  mutate(condition = fct_recode(condition, 
                                'Dementia' = 'dementia',
                                'Peripheral vascular disease' = 'pvd',
                                'Skin ulcers' = 'ulcers',
                                'Congestive heart failure' = 'chf',
                                'Depression' = 'depression',
                                'Diabetes Mellitus' = 'diabetes',
                                'No comorbidity' = 'healthy',
                                # 'Depression' = 'depression',
                                # 'Congestive\nHeart Failure' = 'chf'
                                )) 

d <- conditions %>% select(hrr, smr3, prop, condition) %>% 
  pivot_wider(id_cols = hrr:smr3, names_from='condition', 
              values_from = 'prop')

corr_obs <- d %>% 
  pivot_longer(Dementia:Depression, names_to = 'condition', values_to = 'prop') %>% 
  group_by(condition) %>% 
  summarize(corr_coef = cor(smr3, prop, method = 'spearman')) %>% 
  ungroup()

set.seed(295789)
boot_samples <- do(5000, parallel=FALSE)*
  (resample(d) %>% 
     pivot_longer(cols = Dementia:Depression, 
                  names_to = 'condition',
                  values_to = 'prop') %>% 
     group_by(condition) %>% 
     summarize(corr_coef = cor(smr3, prop, method='spearman')) %>% 
     ungroup())

cor_boot <- boot_samples %>% 
  group_by(condition) %>% 
  summarize(lcb = quantile(corr_coef, 0.025),
            ucb = quantile(corr_coef, 0.975)) %>% 
  ungroup()
out <-  left_join(corr_obs, cor_boot) %>% 
  mutate_at(vars(-condition), round, 3) %>% 
  rename("Condition" = 'condition', 
         "Spearman correlation" = 'corr_coef', 
         'LCB'='lcb',
         'UCB'='ucb') %>% 
  slice(c(2,7, 6, 1, 3, 4, 5))
kable(out, align='lccc') %>% 
  kable_styling(full_width = T)

```

